Himalayan Expeditions Success Analysis

### 2024-04-18 paper/paper.qmd
Plots for data section added

(Plot the variables)

@@

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-results
#| fig-cap: Results1

ggplot(predictions, aes(x = sex, y = success_probability, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Predicted Success Probability by Gender",
       x = "Sex",
       y = "Success Probability") +
  theme_minimal()

@@

# Discussion {#sec-discussion}
1) What is done in this paper? 
2) What is something that we learn about the world?

>>>>

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Counts for the variables of interest
#| fig-subcap: ["Sex", "Solo Status", "Age", "Seasons", "Height of the Peak"]
#| label: fig-counts
#| layout-ncol: 2

plot_bar <- function(data, x, x_label) {
  if (x == "age_range") {
    custom_order <- c("Under 18", "19-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90")
    data[[x]] <- factor(data[[x]], levels = custom_order)
  }

  ggplot(data, aes_string(x = x)) +
    geom_bar(fill = "lightblue") +
    labs(x = x_label, y = "Count") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

plot_bar(analysis_data, "sex", "Sex")
plot_bar(analysis_data, "solo", "Solo Status")
plot_bar(analysis_data, "age_range", "Age")
plot_bar(analysis_data, "seasons", "Season")
plot_bar(analysis_data, "height_range", "Height of the Peak")

@fig-counts shows

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Proportions for the variables of interest and the outcome of their expedition
#| fig-subcap: ["Sex", "Solo Status", "Age", "Seasons", "Height of the Peak"]
#| label: fig-proportions
#| layout-ncol: 2

library(tidyr)

# Define a function to create the plot for each variable
create_plot <- function(data, x_var, x_label) {
  custom_order <- c("Under 18", "19-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90")

  # Create a dataframe with all possible combinations of x_var, success, and died
  all_combinations <- expand.grid(
    x_var = unique(data[[x_var]]),
    success = c(TRUE, FALSE),
    died = c(TRUE, FALSE)
  )

  # Join the original data with all_combinations to fill in missing combinations
  plot_data <- all_combinations %>%
    left_join(data, by = c(x_var = x_var, success = "success", died = "died")) %>%
    group_by(x_var, success, died) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(x_var) %>%
    mutate(proportion = count / sum(count)) %>%
    filter(success | died)

  # If x_var is "age_range", convert it to a factor with custom order
  if (x_var == "age_range") {
    plot_data$x_var <- factor(plot_data$x_var, levels = custom_order)
  }

  p <- ggplot(plot_data, aes(x = x_var, y = proportion, fill = factor(success))) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_fill_manual(values = c("TRUE" = "green", "FALSE" = "red"), 
                      labels = c("Success", "Died"),
                      name = "Outcome") +
    labs(x = x_label, y = "Proportion") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  return(p)
}

# Create plots for each variable
create_plot(analysis_data, "sex", "Sex")
create_plot(analysis_data, "solo", "Solo Status")
create_plot(analysis_data, "age_range", "Age")
create_plot(analysis_data, "seasons", "Season")
create_plot(analysis_data, "height_range", "Height of the Peak")
```

@fig-proportions shows

@@

\newpage

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-sex
#| fig-cap: Predicted Success Probability by Sex

ggplot(predictions, aes(x = sex, y = success_probability, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Predicted Success Probability by Sex",
       x = "Sex",
       y = "Success Probability") +
  theme_minimal()
```
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-age
#| fig-cap: Predicted Success Probability by Age

# Define custom order for age_range levels
custom_order <- c("Under 18", "19-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90")

# Convert age_range to factor with custom order
predictions$age_range <- factor(predictions$age_range, levels = custom_order)

ggplot(predictions, aes(x = age_range, y = success_probability, fill = age_range)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Predicted Success Probability by Age",
       x = "Age",
       y = "Success Probability") +
  theme_minimal()

```
```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-subplots
#| fig-cap: Predicted Success Probability by Age and Sex

# Define custom order for age_range levels
custom_order <- c("Under 18", "19-30", "31-40", "41-50", "51-60", "61-70", "71-80", "81-90")

# Convert age_range to factor with custom order
predictions$age_range <- factor(predictions$age_range, levels = custom_order)

ggplot(predictions, aes(x = age_range, y = success_probability, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Predicted Success Probability by Age and Sex",
       x = "Age",
       y = "Success Probability") +
  facet_grid(sex ~ .) +  # Create subplots for each sex
  theme_minimal()

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-subplots-height
#| fig-cap: Predicted Success Probability by Height Range and Seasons

ggplot(predictions, aes(x = height_range, y = success_probability, fill = seasons)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Predicted Success Probability by Height Range and Seasons",
       x = "Height Range",
       y = "Success Probability") +
  facet_grid(seasons ~ .) +  # Create subplots for each season
  theme_minimal()

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-subplots-sex-solo
#| fig-cap: Predicted Success Probability by Sex and Solo Status

ggplot(predictions, aes(x = sex, y = success_probability, fill = solo)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Predicted Success Probability by Sex and Solo Status",
       x = "Sex",
       y = "Success Probability") +
  facet_grid(solo ~ .) +  # Create subplots for each solo status
  theme_minimal()
```
# Discussion {#sec-discussion}
1) What is done in this paper? 
2) What is something that we learn about the world?
